---
title: "Supplementary Information: Zoogeography"
output:
  html_document:
    toc: TRUE
    code_download: TRUE
    self_contained: TRUE
---

```{r setup, include=FALSE, cache=FALSE}
library(mgcv)
library(raster)
library(dplyr)
library(stringi)
library(parallel)
library(purrr)
library(ggplot2)
library(viridis)
library(knitr)
library(svglite)
library(sp)
library(readr)
library(rgdal)
library(stringr)
library(classInt)
library(maptools)
library(tibble)
library(mgcv)
library(tidyr)
library(DT)
P <- rprojroot::find_rstudio_root_file
source(P("R/model_reduction.R"))
source(P("R/fit_gam.R"))
source(P("R/relative_contributions.R"))
source(P("R/cross_validation.R"))
source(P("R/cv_gam_by.R"))
set.seed(0)
opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, autodep=TRUE)
```


```{r db-loading, cache = TRUE}
db <- readRDS(P("intermediates", "postprocessed_database.rds"))
hosts <- db[["hosts"]]
host_names <- read_csv(P("data/hosts.csv")) %>%
  dplyr::select(hHostNameFinal, Common_names_Eng)

```
# Zoogeographic Cross-Validation

In addition to randomly-selected k-fold cross-validation, we evaluated the robustness of our models via non-random geographic cross-validation: we systematically removed all observations from a zoogeographical region, re-fit the model using all observations from outside the region, then performed a non-parametric permutation test comparing the predicted values to the observed values for that region.


```{r map-setup, cache = TRUE, message=FALSE}
zregions <- shapefile(P("shapefiles", "mam", "mam.shp"), verbose = FALSE)
#Intersected with IUCN habitat shapefile in QGIS
zoog_mammals <- shapefile(P("shapefiles", "host_zg_area", "host_zg_area.shp"), verbose = FALSE)
zoog_mammals_data <- as_tibble(zoog_mammals@data)

#Paring down the dataset
zgm_reduced <- zoog_mammals_data %>%
  select(BINOMIAL, mam_upgma_, garea2)
#Assigning only largest areas
zgm <- zgm_reduced %>%
  group_by(BINOMIAL) %>%
  mutate(area = max(garea2)) %>%
  filter(garea2 == area) %>%
  arrange(BINOMIAL)
zgm <- unique(zgm)

region_names <- readRDS(P("data/region_names.rds"))

sp_rng_lookup <- left_join(zgm, region_names, by = c("mam_upgma_" = "fold")) %>%
  select(name, BINOMIAL) %>%
  group_by(name) %>%
  nest()



```

## Zoonoses GAM - All Associations

```{r all-zoo, cache=TRUE}
data_set = hosts %>% 
  filter(hMarOTerr == "Terrestrial",
         hWildDomFAO == "wild",
         !is.na(PdHoSa.cbCst_order))

#  Create dummy variables for orders to use as random effects
dummys = as.data.frame(with(data_set, model.matrix(~hOrder))[,-1])
data_set = cbind(data_set, dummys)

###### ALL ZOONOSES
allz_gam = readRDS(P("intermediates", "all_zoonoses_models.rds"))$model[[1]]

joined_dat <- left_join(data_set, zgm, by=c("hHostNameFinal"="BINOMIAL"))
joined_dat <- filter_down(allz_gam, joined_dat, "mam_upgma_")

results<-cv_gam_by_zg(allz_gam, joined_dat,unique(joined_dat$mam_upgma_))

allz_dat <- joined_dat

##############

zregions@data <- left_join(zregions@data, results, by=c("mam_upgma_"="fold"))

data("wrld_simpl")
small_world <- subset(wrld_simpl, wrld_simpl@data$NAME != "Antarctica")
plot(small_world)

zregions %>%
  subset(zregions@data$p_value < 0.05 & zregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="red")

zregions %>%
  subset(zregions@data$p_value < 0.05 & zregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="pink")

zregions %>%
  subset(zregions@data$p_value >= 0.05 & zregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="darkgreen")

zregions %>%
  subset(zregions@data$p_value >= 0.05 & zregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="lightgreen")

title("All Zoonoses: 31 Regions")
```

Dark green indicates unbiased regions, while dark red indicates regions with evidence of biased predictions. Light green and light red represent the same distinction, but these regions contain very few (less than ten) assigned species; blank areas were not assigned any hosts. 

```{r all-zoo-tables, results = "asis"}
results<-left_join(results, region_names, by = "fold")

results %>% arrange(fold) %>%
  filter(n_validate > 10, p_value < 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Zoonoses)` = mean_diff) %>%
  kable(caption = "Biased Predictions Regions (n > 10)", digits = 4)

results %>% arrange(fold) %>%
  filter(n_validate > 10, p_value >= 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Zoonoses)` = mean_diff) %>%
  kable(caption = "Unbiased Prediction Regions (n > 10)", digits = 4)

```


## All Viruses GAM - All Associations

```{r all-viruses, cache = TRUE}
###### ALL VIRUSES
allv_gam = readRDS(P("intermediates", "all_viruses_models.rds"))$model[[1]]
vregions <- shapefile(P("shapefiles", "mam", "mam.shp"), verbose = FALSE)

temp <- left_join(data_set, zgm, by=c("hHostNameFinal"="BINOMIAL"))
temp <- filter_down(allv_gam, temp, "mam_upgma_")



allv_dat <- temp

vresults<-cv_gam_by_zg(allv_gam, temp,unique(temp$mam_upgma_))

vregions@data <- left_join(vregions@data, vresults, by=c("mam_upgma_"="fold"))

plot(small_world)

vregions %>%
  subset(vregions@data$p_value < 0.05 & vregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="red")

vregions %>%
  subset(vregions@data$p_value < 0.05 & vregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="pink")

vregions %>%
  subset(vregions@data$p_value >= 0.05 & vregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="darkgreen")

vregions %>%
  subset(vregions@data$p_value >= 0.05 & vregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="lightgreen")

title("All Viruses: 31 Regions")
```

Dark green indicates unbiased regions, while dark red indicates regions with evidence of biased predictions. Light green and light red represent the same distinction, but these regions contain very few (less than ten) assigned species; blank areas were not assigned any hosts. 

```{r all-virus-tables, results = "asis"}
vresults<-left_join(vresults, region_names, by = "fold")

vresults %>% arrange(fold) %>%
  filter(n_validate > 10, p_value < 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Viruses)` = mean_diff) %>%
  kable(caption = "Biased Predictions Regions (n > 10)", digits = 4)

vresults %>% arrange(fold) %>%
  filter(n_validate > 10, p_value >= 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Viruses)` = mean_diff) %>%
  kable(caption = "Unbiased Prediction Regions (n > 10)", digits = 4)

```



## Zoonoses GAM - Strict Associations

```{r strict-zoo, cache = TRUE}
strictz_gam <- readRDS(P("intermediates", "all_zoonoses_strict_models.rds"))$model[[1]]

szregions <- shapefile(P("shapefiles", "mam", "mam.shp"), verbose = FALSE)

temp <- left_join(data_set, zgm, by=c("hHostNameFinal"="BINOMIAL"))
temp <- filter_down(strictz_gam, temp, "mam_upgma_")

strictz_dat <- temp

#########

szresults<-cv_gam_by_zg(strictz_gam, temp,unique(temp$mam_upgma_))

szregions@data <- left_join(szregions@data, szresults, by=c("mam_upgma_"="fold"))

plot(small_world)

szregions %>%
  subset(szregions@data$p_value < 0.05 & szregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="red")

szregions %>%
  subset(szregions@data$p_value < 0.05 & szregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="pink")

szregions %>%
  subset(szregions@data$p_value >= 0.05 & szregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="darkgreen")

szregions %>%
  subset(szregions@data$p_value >= 0.05 & szregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="lightgreen")

title("Strict Zoonoses: 31 Regions")
```

Dark green indicates unbiased regions, while dark red indicates regions with evidence of biased predictions. Light green and light red represent the same distinction, but these regions contain very few (less than ten) assigned 
species; blank areas were not assigned any hosts.  

```{r sz-tables, results = "asis"}
szresults<-left_join(szresults, region_names, by = "fold")

szresults %>% arrange(fold) %>%
  filter(n_validate > 10, p_value < 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Zoonoses)` = mean_diff) %>%
  kable(caption = "Biased Predictions Regions (n > 10)", digits = 4)
```

```{r sz-tables2}
szresults %>% arrange(fold) %>%
  filter(n_validate > 10, p_value >= 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Zoonoses)` = mean_diff) %>%
  kable(caption = "Unbiased Prediction Regions (n > 10)", digits = 4)

```


## All Viruses GAM - Strict Associations

```{r strict-viruses, cache = TRUE}
strictv_gam <- readRDS(P("intermediates", "all_viruses_strict_models.rds"))$model[[1]]

svregions <- shapefile(P("shapefiles", "mam", "mam.shp"), verbose = FALSE)

temp <- left_join(data_set, zgm, by=c("hHostNameFinal"="BINOMIAL"))
temp <- filter_down(strictv_gam, temp, "mam_upgma_")

strictv_dat <- temp

svresults<-cv_gam_by_zg(strictv_gam, temp,unique(temp$mam_upgma_))

svregions@data <- left_join(svregions@data, svresults, by=c("mam_upgma_"="fold"))

plot(small_world)

svregions %>%
  subset(svregions@data$p_value < 0.05 & svregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="red")

svregions %>%
  subset(svregions@data$p_value < 0.05 & svregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="pink")

svregions %>%
  subset(svregions@data$p_value >= 0.05 & svregions@data$n_validate > 10) %>%
  plot(.,add=TRUE, col="darkgreen")

svregions %>%
  subset(svregions@data$p_value >= 0.05 & svregions@data$n_validate < 10) %>%
  plot(.,add=TRUE, col="lightgreen")

title("Strict Viruses: 31 Regions")
```

Dark green indicates unbiased regions, while dark red indicates regions with evidence of biased predictions. Light green and light red represent the same distinction, but these regions contain very few (less than ten) assigned species; blank areas were not assigned any hosts. 

```{r strict-zoo-tables, results = "asis"}
svresults<-left_join(svresults, region_names, by = "fold")

svresults %>% arrange(fold) %>%
  filter(n_validate > 10, p_value < 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Viruses)` = mean_diff) %>%
  kable(caption = "Biased Predictions Regions (n > 10)", digits = 4)

svresults %>% arrange(fold) %>%
  filter(n_validate > 10, p_value >= 0.05) %>%
  select(name, n_fit, n_validate, p_value, mean_diff) %>%
  rename(Region = name, `Observations Fit` = n_fit, `Observations Held Out` = n_validate, 
         `P-value` = p_value, `Mean Prediction Difference (Number of Viruses)` = mean_diff) %>%
  kable(caption = "Unbiased Prediction Regions (n > 10)", digits = 4)
```


#Region as Random Effect

To further investigate the relationship between zoogeographical region and our outcomes of interest, we added zoographical region as a categorical random effect to each of our best-fit GAMs. The model fit and variable relative deviance explained comparisons for each model follow:

## All Zoonoses GAM
Adding zoogeographical region as a categorical random effect to our best-fit All Zoonoses GAM does not improve the model fit, as seen in the tables below:

```{r non-sig-re}
#All zoonoses
allz_dat <- allz_dat %>%
  mutate(zg_region = as.factor(mam_upgma_))

allz_region_gam <- gam(NSharedWithHoSa ~ s(hDiseaseZACitesLn, bs = "tp", k = 7) + s(hHuntedIUCN, bs = "re") + s(hOrderCETARTIODACTYLA, bs = "re") + 
                       s(hOrderCHIROPTERA, bs = "re") + s(hOrderDIPROTODONTIA, bs = "re") + 
                       s(hOrderPERAMELEMORPHIA, bs = "re") + s(hOrderPERISSODACTYLA, bs = "re") + s(zg_region, bs = "re") +
                       s(hOrderSCANDENTIA, bs = "re") + s(PdHoSa.cbCst, bs = "tp", k = 7) + 
                       s(UrbRurPopRatioLn, bs = "tp", k = 7) + offset(LnTotNumVirus), poisson, allz_dat, select=TRUE)

allz_rep_gam <- gam(NSharedWithHoSa ~ s(hDiseaseZACitesLn, bs = "tp", k = 7) + s(hHuntedIUCN, bs = "re") + s(hOrderCETARTIODACTYLA, bs = "re") + 
                       s(hOrderCHIROPTERA, bs = "re") + s(hOrderDIPROTODONTIA, bs = "re") + 
                       s(hOrderPERAMELEMORPHIA, bs = "re") + s(hOrderPERISSODACTYLA, bs = "re") +
                       s(hOrderSCANDENTIA, bs = "re") + s(PdHoSa.cbCst, bs = "tp", k = 7) + 
                       s(UrbRurPopRatioLn, bs = "tp", k = 7) + offset(LnTotNumVirus), poisson, allz_dat, select=TRUE)

AIC <- c(allz_rep_gam$aic, allz_region_gam$aic)
`Deviance Explained` <- c(1-allz_rep_gam$deviance/allz_rep_gam$null.deviance, 1-allz_region_gam$deviance/allz_region_gam$null.deviance )

mc <- t(tibble(AIC,`Deviance Explained`))
colnames(mc) <- c("Original Model", "Region Model")

kable(mc, caption = "All Zoonoses Model Comparison", digits =4)

allz_relo <- get_relative_contribs(allz_rep_gam) %>% 
  rename(original_rel = rel_deviance_explained)

allz_relg <- get_relative_contribs(allz_region_gam) %>% 
  rename(region_rel = rel_deviance_explained)
  
allz_com <- allz_relo %>%
  full_join(allz_relg, by = "term") 

allz_com %>% rename(`Term` = term, `Original Model Relative Deviance Explained` = original_rel, `Region Model Relative Deviance Explained` = region_rel) %>%
  kable(caption = "All Zoonoses: Original vs. Region Model Relative Deviance Explained", digits = 4)

```
  
  
## All Viruses GAM  
Adding zoogeographical region as a categorical random effect to our best-fit All Viruses GAM improves the model, as seen by the decrease in AIC and the increase in deviance explained.

```{r allv-region}
#All viruses
allv_dat <- allv_dat %>%
  mutate(zg_region = as.factor(mam_upgma_))

allv_rep_gam <-gam(TotVirusPerHost ~  s(hDiseaseZACitesLn, bs = "cs", k = 7) + s(hMassGramsPVR, bs = "cs", k = 7) + s(hOrderCETARTIODACTYLA, bs = "re") + 
    s(hOrderCHIROPTERA, bs = "re") + s(hOrderEULIPOTYPHLA, bs = "re") + 
    s(hOrderPERISSODACTYLA, bs = "re") + s(hOrderPRIMATES, bs = "re") + 
    s(hOrderRODENTIA, bs = "re") + s(LnAreaHost, bs = "cs", k = 7) + 
    s(S20, bs = "cs", k = 7), poisson, allv_dat, select = TRUE)

allv_region_gam <- gam(TotVirusPerHost ~  s(hDiseaseZACitesLn, bs = "cs", k = 7) + s(hMassGramsPVR, bs = "cs", k = 7) + s(hOrderCETARTIODACTYLA, bs = "re") + 
    s(hOrderCHIROPTERA, bs = "re") + s(hOrderEULIPOTYPHLA, bs = "re") + 
    s(hOrderPERISSODACTYLA, bs = "re") + s(hOrderPRIMATES, bs = "re") + 
    s(hOrderRODENTIA, bs = "re") + s(LnAreaHost, bs = "cs", k = 7) + s(zg_region, bs = "re") +
    s(S20, bs = "cs", k = 7), poisson, allv_dat, select = TRUE)

AIC <- c(allv_rep_gam$aic, allv_region_gam$aic)
`Deviance Explained` <- c(1-allv_rep_gam$deviance/allv_rep_gam$null.deviance, 1-allv_region_gam$deviance/allv_region_gam$null.deviance )

mc <- t(tibble(AIC,`Deviance Explained`))
colnames(mc) <- c("Original Model", "Region Model")

kable(mc, caption = "All Viruses Model Comparison", digits =4)

allv_relo <- get_relative_contribs(allv_rep_gam) %>% 
  rename(original_rel = rel_deviance_explained)

allv_relg <- get_relative_contribs(allv_region_gam) %>% 
  rename(region_rel = rel_deviance_explained)
  
allv_com <- allv_relo %>%
  full_join(allv_relg, by = "term") 

allv_com %>% rename(`Term` = term, `Original Model Relative Deviance Explained` = original_rel, `Region Model Relative Deviance Explained` = region_rel) %>%
  kable(caption = "All Viruses: Original vs. Region Model Relative Deviance Explained", digits = 4)


```
  

## Strict Zoonoses GAM
Adding zoogeographical region as a categorical random effect to our best-fit Strict Zoonoses GAM improves the model, as seen by the decrease in AIC and the increase in deviance explained.

```{r strictz-region}
# Strict Zoonoses
strictz_dat <- strictz_dat %>%
  mutate(zg_region = as.factor(mam_upgma_))

sz_rep_gam <- gam(NSharedWithHoSa_strict ~ s(hDiseaseZACitesLn, bs = "tp", k = 7) + 
    s(hMassGramsPVR, bs = "tp", k = 7) + s(hOrderCETARTIODACTYLA, 
    bs = "re") + s(hOrderDIPROTODONTIA, bs = "re") + s(hOrderLAGOMORPHA, 
    bs = "re") + s(hOrderPERISSODACTYLA, bs = "re") + s(hOrderPRIMATES, 
    bs = "re") + s(HumPopDensLnChg, bs = "tp", k = 7) + s(PdHoSa.cbCst, 
    bs = "tp", k = 7) + s(UrbRurPopRatioChg, bs = "tp", k = 7) + 
    offset(LnTotNumVirus), poisson, strictz_dat, select = TRUE)

sz_region_gam <- gam(NSharedWithHoSa_strict ~ s(hDiseaseZACitesLn, bs = "tp", k = 7) + 
    s(hMassGramsPVR, bs = "tp", k = 7) + s(hOrderCETARTIODACTYLA, bs = "re") + 
      s(hOrderDIPROTODONTIA, bs = "re") + s(hOrderLAGOMORPHA, bs = "re") + s(zg_region, bs = "re") +
      s(hOrderPERISSODACTYLA, bs = "re") + s(hOrderPRIMATES, bs = "re") + 
      s(HumPopDensLnChg, bs = "tp", k = 7) + s(PdHoSa.cbCst, bs = "tp", k = 7) + 
      s(UrbRurPopRatioChg, bs = "tp", k = 7) + offset(LnTotNumVirus), poisson, strictz_dat, select = TRUE)

AIC <- c(sz_rep_gam$aic, sz_region_gam$aic)
`Deviance Explained` <- c(1-sz_rep_gam$deviance/sz_rep_gam$null.deviance, 1-sz_region_gam$deviance/sz_region_gam$null.deviance )

mc <- t(tibble(AIC,`Deviance Explained`))
colnames(mc) <- c("Original Model", "Region Model")

kable(mc, caption = "Strict Zoonoses Model Comparison", digits =4)

sz_relo <- get_relative_contribs(sz_rep_gam) %>% 
  rename(original_rel = rel_deviance_explained)

sz_relg <- get_relative_contribs(sz_region_gam) %>% 
  rename(region_rel = rel_deviance_explained)
  
sz_com <- sz_relo %>%
  full_join(sz_relg, by = "term") 

sz_com %>% rename(`Term` = term, `Original Model Relative Deviance Explained` = original_rel, `Region Model Relative Deviance Explained` = region_rel) %>%
  kable(caption = "Strict Zoonoses: Original vs. Region Model Relative Deviance Explained", digits = 4)


```
  

##Strict Viruses GAM
Adding zoogeographical region as a categorical random effect to our best-fit Strict Viruses GAM improves the model, as seen by the decrease in AIC and the increase in deviance explained.

```{r strictv-region}
strictv_dat <- strictv_dat %>%
  mutate(zg_region = as.factor(mam_upgma_))

strictv_rep_gam <- gam(TotVirusPerHost_strict ~ s(hDiseaseZACitesLn, bs = "cs", k = 7) + 
    s(hMassGramsPVR, bs = "cs", k = 7) + s(hOrderCHIROPTERA, bs = "re") + 
    s(hOrderCINGULATA, bs = "re") + s(hOrderEULIPOTYPHLA, bs = "re") + 
    s(hOrderPERAMELEMORPHIA, bs = "re") + s(hOrderPRIMATES, bs = "re") + 
    s(hOrderRODENTIA, bs = "re") + s(hOrderSCANDENTIA, bs = "re") + 
    s(LnAreaHost, bs = "cs", k = 7) + s(S20, bs = "cs", k = 7), poisson, strictv_dat, select=TRUE)

strictv_region_gam <- gam(TotVirusPerHost_strict ~ s(hDiseaseZACitesLn, bs = "cs", k = 7) + 
    s(hMassGramsPVR, bs = "cs", k = 7) + s(hOrderCHIROPTERA, bs = "re") + 
    s(hOrderCINGULATA, bs = "re") + s(hOrderEULIPOTYPHLA, bs = "re") + 
    s(hOrderPERAMELEMORPHIA, bs = "re") + s(hOrderPRIMATES, bs = "re") + s(zg_region, bs = "re") +
    s(hOrderRODENTIA, bs = "re") + s(hOrderSCANDENTIA, bs = "re") + 
    s(LnAreaHost, bs = "cs", k = 7) + s(S20, bs = "cs", k = 7), poisson, strictv_dat, select=TRUE)


AIC <- c(strictv_rep_gam$aic, strictv_region_gam$aic)
`Deviance Explained` <- c(1-strictv_rep_gam$deviance/strictv_rep_gam$null.deviance, 1-strictv_region_gam$deviance/strictv_region_gam$null.deviance )

mc <- t(tibble(AIC,`Deviance Explained`))
colnames(mc) <- c("Original Model", "Region Model")

kable(mc, caption = "Strict Viruses Model Comparison", digits =4)

strictv_relo <- get_relative_contribs(strictv_rep_gam) %>% 
  rename(original_rel = rel_deviance_explained)

strictv_relg <- get_relative_contribs(strictv_region_gam) %>% 
  rename(region_rel = rel_deviance_explained)
  
strictv_com <- strictv_relo %>%
  full_join(strictv_relg, by = "term") 

strictv_com %>% rename(`Term` = term, `Original Model Relative Deviance Explained` = original_rel, `Region Model Relative Deviance Explained` = region_rel) %>%
  kable(caption = "Strict Viruses: Original vs. Region Model Relative Deviance Explained", digits = 4)

```
  

