---
title: 'Supplementary Information: GAM Models'
output:
  html_document:
    toc: TRUE
    code_download: TRUE
    self_contained: TRUE
---

```{r setup, include=FALSE, cache=FALSE}
library(mgcv)
library(dplyr)
library(stringi)
library(parallel)
library(purrr)
library(ggplot2)
library(viridis)
library(knitr)
library(svglite)
P <- rprojroot::find_rstudio_root_file
source(P("R/model_reduction.R"))
source(P("R/fit_gam.R"))
source(P("R/relative_contributions.R"))
source(P("R/cross_validation.R"))
set.seed(0)
opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, autodep=TRUE)
```

# Zoonoses GAM - All Associations

```{r get-all-zoo, cache=FALSE}
all_zoonoses <- readRDS(P("intermediates", "all_zoonoses_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r all-zoo-table, cache=FALSE}
all_zoonoses %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r all-zoo-plot, fig.width=10, fig.height=12, cache=FALSE}
plot(all_zoonoses$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r all-zoo-summary, cache=FALSE}
summary(all_zoonoses$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r all-zoo-contrib, cache=FALSE}
get_relative_contribs(all_zoonoses$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r all-zoo-cv, cache=TRUE}
cv_gam(all_zoonoses$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# Zoonoses GAM - Strict Associations

```{r get-all-zoo-strict, cache=TRUE}
all_zoonoses_strict <- readRDS(P("intermediates", "all_zoonoses_strict_models.rds"))

```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r all-zoo-strict-table, cache=FALSE}
all_zoonoses_strict %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r all-zoo-strict-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(all_zoonoses_strict$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r all-zoo-strict-summary, cache=FALSE}
summary(all_zoonoses_strict$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r all-zoo-strict-contrib, cache=FALSE}
get_relative_contribs(all_zoonoses_strict$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r all-zoo-strict-cv, cache=TRUE}
cv_gam(all_zoonoses_strict$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# Zoonoses GAM - All Associations without Reverse Zoonoses

```{r fit-all-zoo-norev, cache=TRUE}
all_zoonoses_norev <- readRDS(P("intermediates", "all_zoonoses_norev_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r all-zoo-norev-table, cache=FALSE}
all_zoonoses_norev %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r all-zoo-norev-plot, fig.width=10, fig.height=12, cache=FALSE}
plot(all_zoonoses_norev$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r all-zoo-norev-summary, cache=FALSE}
summary(all_zoonoses_norev$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r all-zoo-norev-contrib, cache=FALSE}
get_relative_contribs(all_zoonoses_norev$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r all-zoo-norev-cv, cache=TRUE}
cv_gam(all_zoonoses_norev$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# All Viruses GAM - All Associations

```{r fit-all-vir, cache=TRUE}
all_viruses <- readRDS(P("intermediates", "all_viruses_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r all-vir-table, cache=FALSE}
all_viruses %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r all-vir-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(all_viruses$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```


Summary of top model:

```{r all-vir-summary, cache=FALSE}
summary(all_viruses$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r all-vir-contrib, cache=FALSE}
get_relative_contribs(all_viruses$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r all-vir-cv, cache=TRUE}
cv_gam(all_viruses$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# All Viruses GAM - Strict Associations

```{r all-vir-strict-get, cache=TRUE}
all_viruses_strict <- readRDS(P("intermediates", "all_viruses_strict_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r all-vir-strict-table, cache=FALSE}
all_viruses_strict %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r all-vir-strict-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(all_viruses_strict$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r all-vir-strict-summary, cache=FALSE}
summary(all_viruses_strict$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r all-vir-strict-contrib, cache=FALSE}
get_relative_contribs(all_viruses_strict$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r all-vir-stict-cv, cache=TRUE}
cv_gam(all_viruses_strict$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# Viral Traits GAM - All Associations

```{r vtraits-get, cache=TRUE}
vtraits <- readRDS(P("intermediates", "vtraits_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r vtraits-table, cache=FALSE}
vtraits %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r vtraits-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(vtraits$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r vtraits-summary, cache=FALSE}
summary(vtraits$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r vtraits-contrib, cache=FALSE}
get_relative_contribs(vtraits$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r vtraits-cv, cache=TRUE}
cv_gam(vtraits$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```


# Viral Traits GAM - Strict Associations


```{r vtraits-strict-get, cache=TRUE}
vtraits_strict <- readRDS(P("intermediates", "vtraits_strict_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r vtraits-strict-table, cache=FALSE}
vtraits_strict %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r vtraits-strict-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(vtraits_strict$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r vtraits-strict-summary, cache=FALSE}
summary(vtraits_strict$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r vtraits-strict-contrib, cache=FALSE}
get_relative_contribs(vtraits_strict$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r vtraits-strict-cv, cache=TRUE}
cv_gam(vtraits_strict$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# Zoonoses in Domestic Animals GAM - All Associations

```{r dom-get, cache=TRUE}
domestic_zoonoses <- readRDS(P("intermediates", "domestic_zoonoses_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r dom-table, cache=FALSE}

domestic_zoonoses %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r dom-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(domestic_zoonoses$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r dom-summary, cache=FALSE}
summary(domestic_zoonoses$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r dom-contrib, cache=FALSE}
get_relative_contribs(domestic_zoonoses$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r dom-cv, cache=TRUE}
cv_gam(domestic_zoonoses$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```


# Zoonoses in Domestic Animals GAM - Strict Associations

```{r dom-strict-get, cache=TRUE}
domestic_zoonoses_strict <- readRDS(P("intermediates", "domestic_zoonoses_strict_models.rds"))
```

Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r dom-strict-table, cache=FALSE}
domestic_zoonoses_strict %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r dom-strict-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(domestic_zoonoses_strict$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r dom-strict-summary, cache=FALSE}
summary(domestic_zoonoses_strict$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r dom-strict-contrib, cache=FALSE}
get_relative_contribs(domestic_zoonoses_strict$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r dom-strict-cv, cache=TRUE}
cv_gam(domestic_zoonoses_strict$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# All Viruses in Domestic Animals GAM - All Associations

```{r dom-vir-get, cache=FALSE}
domestic_viruses <- readRDS(P("intermediates", "domestic_viruses_models.rds"))
```


Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r dom-vir-table, cache=FALSE}
model_family = poisson
domestic_viruses %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model` = terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r dom-vir-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(domestic_viruses$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r dom-vir-summary, cache=FALSE}
summary(domestic_viruses$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r dom-vir-contrib, cache=FALSE}
get_relative_contribs(domestic_viruses$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r dom-vir-cv, cache=TRUE}
cv_gam(domestic_viruses$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```

# All Viruses in Domestic Animals GAM - Stringent Associations

```{r dom-vir-strict-get, cache=TRUE}
domestic_viruses_strict <- readRDS(P("intermediates", "domestic_viruses_strict_models.rds"))
```


Terms in models with ΔAIC < 2.  All continuous terms effects are represented
as splines, all discrete terms as random effects:

```{r dom-vir-strict-table, cache=FALSE}
domestic_viruses_strict %>%
  filter(daic <= 2) %>%
  select(daic, terms) %>% 
    mutate(terms = stri_replace_all_regex(terms, "s\\(([^\\)]*)\\)", "$1")) %>%
  rename(ΔAIC = daic, `Terms in Model`=terms) %>% 
  kable()
```

Partial effect plots of all terms in top model:

```{r dom-vir-strict-plot, fig.width=10, fig.height=10, cache=FALSE}
plot(domestic_viruses_strict$model[[1]], all.terms=TRUE, pages=1, scale=0, residuals = TRUE)
```

Summary of top model:

```{r dom-vir-strict-summary, cache=FALSE}
summary(domestic_viruses_strict$model[[1]])
```

Relative fraction of deviance explained by each variable in the top model:

```{r dom-vir-strict-contrib, cache=FALSE}
get_relative_contribs(domestic_viruses_strict$model[[1]]) %>%
    mutate(rel_deviance_explained = paste0(signif(rel_deviance_explained*100, digits=3), "%")) %>% 
  rename(Term = term, `Relative Deviance Explained` = rel_deviance_explained) %>% 
  kable(align=c("l","r"))
```

10-fold cross-validation.  Good fit indicated by *non*-significant p-values:

```{r dom-vir-strict-cv, cache=TRUE}
cv_gam(domestic_viruses_strict$model[[1]]) %>%
    mutate(p_value = round(p_value, digits=3)) %>% 
  rename(Fold = fold, `Observations Fit`=n_fit, `Observations Held Out` = n_validate,
         `P-value`=p_value, `Mean Error` = mean_error) %>%
  kable(align=c("l","r", "r", "r", "r"))
```
